---
title: "Calculating Probabilistic Dynamic Exposure Concentrations"
author: "Joris T.K. Quik and Carlos Blanco"
date: "20/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

ptm <- proc.time()
```

## Introduction

This document describes the code and functions to produce quasi-dynamic ('levelIV') solutions of the [SimpleBox](https://www.rivm.nl/simplebox) multimedia fate model in comvbination with dynamic emission rates.

This script uses a adaptation of SimpleBox v4 to which a local scale with air, water, sediment and soil compartments. The SimnpleBox model is run probabilistically seperately to produce a series of K matrices which are dynamically analysed using this script in combination with probabilistic dynamic emission data.

```{r requirements}
library(openxlsx)
library(deSolve)
library(ggplot2)
library(reshape2)
library(devtools)
library(foreach)
library(doParallel)
# adjust to correct location of SimpleBox xls file
# sb4n.loc <- paste("data/SimpleBox4.0_web_PBT1.xlsm",sep="")

#For installing correct version of openxlsx:
#install_version("openxlsx", version = "4.2.3", repos = "https://cran.us.r-project.org")

```


## required functions
```{r}


# ODE function of SimpleBox:

SimpleBoxODE <- function(t, m, parms) {
  
  with(as.list(c(parms, m)), {
    e <- c(rep(0,length(SB.names)))
    
    e[grep("aL",SB.names)] <- ef.a_L(t)  
    e[grep("wL",SB.names)] <- ef.fw_L(t)  
    e[grep("sL",SB.names)] <- ef.s_L(t)
    
    e[grep("aR",SB.names)] <- ef.a_R(t)  
    e[grep("w1R",SB.names)] <- ef.fw_R(t)  
    e[grep("s2R",SB.names)] <- ef.s_R(t)  # soil emission goes to agricultural soil
    
    e[grep("aC",SB.names)] <- ef.a_C(t)  
    e[grep("w1C",SB.names)] <- ef.fw_C(t)  
    e[grep("s2C",SB.names)] <- ef.s_C(t)  # soil emission goes to agricultural soil
    
    dm <- K %*% m + e
    res <- c(dm)
    list(res, signal = e)
    
  })
}

```


## Probabilistic Dynamic analysis
### inputs

```{r inputs}
# base SimpleBox file
# sb4n.loc <- "data/SimpleBox4DIRECT 07102021.xlsm"
sb4n.loc <- "data/@RISK_SimpleBox4DIRECT 28012022.xlsm"

# SB.K <- as.matrix(read.xlsx(sb4n.loc,colNames=FALSE, namedRegion ="K")) # matrix of rate constants "k"
#SB.m0 <- read.xlsx(sb4n.loc,colNames=FALSE, namedRegion ="m0") # Initial mass of each compartment
SB.names <- read.xlsx(sb4n.loc,colNames=FALSE, namedRegion ="box_names") #Names for each compartment
# SB.v <- read.xlsx(sb4n.loc,colNames=FALSE, namedRegion ="v") #Volumes

# start with 0 mass of substance x
SB.m0 <- rep(0,length(SB.names)) # Initial mass of each compartment in kg

# Read in and prepare emission data per compartment and scale
Analysis_name <- "v0.9"
# Scenario_name <- "III-V_Si_As_NoR"

detectCores()
ncores <- 12 #take care not to set to more than amount of cores available.

# read in emission data

cases <- expand.grid(Recycling=c("NR", "R"),substance=c("Ag","Cu", "Pb"),paneltype=c("Tandem"))
cases$casename <- paste(cases$substance,cases$Recycling,cases$paneltype,sep="_")
```

### main dynamic calculations
```{r main dyn calculation}

out.data2 <- as.list(NULL)

for(i in 1:length(cases[,1])){
  
  ################################################
  # prepare emission data for emission functions #
  
  Emis.Loc <- paste0("data/Emis PV data tandem perovskite/PV_emissions_",cases$paneltype[i],"_v0.1_LOC_",cases$Recycling[i],"_",cases$substance[i],"_kg.xlsm") 
  Emis.Reg <- paste0("data/Emis PV data tandem perovskite/PV_emissions_",cases$paneltype[i],"_v0.1_AM_",cases$Recycling[i],"_",cases$substance[i],"_kg.xlsm") 
  Emis.Cont <-paste0("data/Emis PV data tandem perovskite/PV_emissions_",cases$paneltype[i],"_v0.1_EU_",cases$Recycling[i],"_",cases$substance[i],"_kg.xlsm") 
  
  
  prep.emis.v1 <- function(emis.data){ # input is format of xls emission data
    emis.data[,1] <- paste0(emis.data[,1], Analysis_name)
    row.names(emis.data) <- emis.data[,1]
    emis.data[,1 ] <- 0
    colnames(emis.data)[1] <- 0
    emis.data <- emis.data/(365.25*24*3600)
    emis.data # output is formatted emission data for aproxfun + desolve
  }
  
  # emission rate is read in and organised
  Emis_a_L <- prep.emis.v1(read.xlsx(Emis.Loc,colNames=TRUE,sheet = "PROB_Y_a")) 
  Emis_fw_L <- prep.emis.v1(read.xlsx(Emis.Loc,colNames=TRUE,sheet = "PROB_Y_tw"))
  Emis_s_L <- prep.emis.v1(read.xlsx(Emis.Loc,colNames=TRUE,sheet = "PROB_Y_s")) 
  
  Emis_a_R <- prep.emis.v1(read.xlsx(Emis.Reg,colNames=TRUE,sheet = "PROB_Y_a")) 
  Emis_fw_R <- prep.emis.v1(read.xlsx(Emis.Reg,colNames=TRUE,sheet = "PROB_Y_tw"))
  Emis_s_R <- prep.emis.v1(read.xlsx(Emis.Reg,colNames=TRUE,sheet = "PROB_Y_s")) 
  
  Emis_a_C <- prep.emis.v1(read.xlsx(Emis.Cont,colNames=TRUE,sheet = "PROB_Y_a"))
  Emis_fw_C <- prep.emis.v1(read.xlsx(Emis.Cont,colNames=TRUE,sheet = "PROB_Y_tw"))
  Emis_s_C <- prep.emis.v1(read.xlsx(Emis.Cont,colNames=TRUE,sheet = "PROB_Y_s"))
  
  # Function for converting emission data to linear interpolation functions for Solver
  f.emisfun <- function(Y){
    f.emisfun.a <- function(X){
      approxfun(data.frame(year = as.numeric(names(X))*(60*60*24*365.25),
                           emis_kg = as.numeric(X)),rule = 1:2)
    }
    Y.list <- setNames(split(Y,                 # Modify names of list elements
                             seq(nrow(Y))),
                       rownames(Y))
    emisfun.Y <- lapply(Y.list,f.emisfun.a)
    emisfun.Y
  }
  
  # creates a list of emission data:
  Emisdata <- list( Emis_a_L=Emis_a_L,
                    Emis_fw_L=Emis_fw_L,
                    Emis_s_L=Emis_s_L,
                    Emis_a_R=Emis_a_R,
                    Emis_fw_R=Emis_fw_R,
                    Emis_s_R=Emis_s_R,
                    Emis_a_C=Emis_a_C,
                    Emis_fw_C=Emis_fw_C,
                    Emis_s_C=Emis_s_C)
  Emisfunctions <- lapply(Emisdata,f.emisfun) # converts elements in list to emission functions
  
  ####################################
  # Define Location of K mastrices   #
  
  ADDRISKdata <- paste0("data/SBMatrix C-si/",cases$substance[i],"_SB4DIRECT @RISK output to K-matrix_27012022.xlsx")
  
  ####################################
  # start of parallel computing      #
  registerDoParallel(ncores)  # use multicore, set to the number of our cores
  
  out.data <- 
    foreach(k=c(1:1000), .inorder=FALSE, .errorhandling = "pass") %dopar%  {
      
      Run <- paste0(k, Analysis_name)
      
      ## for reading in K matrices (SB.K) ##
      n <- 0:(999)
      matrixK.rijstart = 2
      matrixK.rij <- matrixK.rijstart+n*39
      matrixK.kollom <- 3  
      
      SB.K = as.matrix(read.xlsx(ADDRISKdata,
                                 colNames=TRUE, 
                                 sheet="Matrixes",
                                 rows = c(matrixK.rij[k]:(matrixK.rij[k]+37)),
                                 cols = c(matrixK.kollom:(matrixK.kollom+37))) )
      
      parms <- list(K=as.matrix(SB.K),SB.names,Run,
                    ef.a_L=Emisfunctions[["Emis_a_L"]][[Run]],
                    ef.fw_L=Emisfunctions[["Emis_fw_L"]][[Run]],
                    ef.s_L=Emisfunctions[["Emis_s_L"]][[Run]],
                    ef.a_R=Emisfunctions[["Emis_a_R"]][[Run]],
                    ef.fw_R=Emisfunctions[["Emis_fw_R"]][[Run]],
                    ef.s_R=Emisfunctions[["Emis_s_R"]][[Run]],
                    ef.a_C=Emisfunctions[["Emis_a_C"]][[Run]],
                    ef.fw_C=Emisfunctions[["Emis_fw_C"]][[Run]],
                    ef.s_C=Emisfunctions[["Emis_s_C"]][[Run]])
      
      etimes <- c(0:100)*(60*60*24*365.25) # manually set output times
      out <- ode(y=as.numeric(SB.m0),times=c(etimes),func=SimpleBoxODE,parms,rtol=1e-30,atol=1e-7)
      colnames(out)[2:38] <- SB.names
      
      list(SBout_kg=out,Run=Run)
      # output to out.data list
      
      # list of all the runs. Each out object contains time step, the mass in each compartment (kg) and a set of signals, which are the emission rates for each compartment.
      
      
    }
  stopImplicitCluster()
  out.data2[[cases$casename[i]]] <- out.data
  addriskinputs <- read.xlsx(ADDRISKdata,
                             colNames=TRUE,
                             rowNames = FALSE,
                             sheet="Raw data output",
                             rows = c(2:1005),
                             cols = c( 1369:1526)) 
  out.data2[[cases$casename[i]]]$addriskinput <- addriskinputs
  
}




# head(out.data2)

# save data for use in R later
save(out.data2, file=paste0("data/20220317RprobDynSB4_perov_tandem.RData"))

# Stop the clock
runtime <- proc.time() - ptm
print(paste(round(runtime[["elapsed"]]/(60),0),"minutes runtime"))


# for testing
# plot(out)
#head(out.data2)
#str(out.data2)
# str(out.data2$Ga_R[[1]]$SBout_kg)
# 
# plot(out.data2[[1]][[1]]$SBout_kg)

# out.data2$Ga_R[[1]]$SBout_kg
# without running all, read in output:

```

