---
title: "Figures for Factsheet Veilige en duurzame zonnepanelen impact voorspellen tijdens ontwerp"
author: "Joris T.K. Quik, Carlos Blanco, and Matthias Hof"
date: "15/08/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Load required data, packages, and fuctions
```{r}
# input data
load("data/20220627RprobDynSB4_perov_tandem.RData")

# required packages
library(tidyr)
library(openxlsx)
library(ggplot2)
library(dplyr)
library(psych)
library(RColorBrewer)

# relevant functions

f_Soil.wetweight <- function(Conc.soil, # in kg/m3 soil or sediment
                             Fracw,
                             Fraca,
                             RHOsolid){
  Conc.soil*1000/(Fracw*1000+(1-Fracw-Fraca)*RHOsolid) # in g/kg (wet) soil
  
}

f_Soil.dryweight <- function(Conc.soil, # in kg/m3 soil
                             Fracs,
                             RHOsolid){
  Conc.soil*1000/(Fracs*RHOsolid) # in g/kg (dry) soil
}

f_wet2dryweight <- function(Conc.soil, # in g/kg (wet) soil/sediment
                            Fracs,
                            RHOsolid,
                            Fracw,
                            RHOw){
  Conc.soil*(Fracw*RHOw/(Fracs*RHOsolid)+1) # in g/kg (dry) soil
}

cases <- expand.grid(Recycling=c("NR", "R"),substance=c("Ag","Cu", "Pb"),paneltype=c("Tandem"))
cases$casename <- paste(cases$substance,cases$Recycling,cases$paneltype,sep="_")
SB.scales <- c("LOC", "AM", "EU")
Substances <- as.character(unique(cases$substance))

```

#### Figure 3. "Emissie in de tijd naar het milieu van lood (Pb) afkomstig van C-Si/Lood-perovskiet tandem zonnepanelen op lokale schaal bij storten (EOL_R) of recycling (EOL_NR) als einde levensfase en emissie door uitloging tijdens gebruik."

```{r figure 3}

#initialize variables

emis.scens <- expand.grid("Scales"=SB.scales,"Substances"=Substances)
Emis_routes <- c("USE_s2", "USE_w0", "USE_w1", "EOL_a_NR", "EOL_a_R", "EOL_s2_NR","EOL_s2_R")

emis.stat.long3 <- data.frame()

## Makes a matrix with concentrations (year x MC run) from deSolve lists, for each compartment and scenario

for(s in 1:length(emis.scens[,1])){
  Emis.loc_NR <- paste0("data/Emis PV data tandem perovskite/PV_emissions_Tandem_v0.1_",
                        emis.scens$Scales[s],"_NR_",
                        emis.scens$Substances[s],"_kg.xlsm") 
  Emis.loc_R <- paste0("data/Emis PV data tandem perovskite/PV_emissions_Tandem_v0.1_",
                       emis.scens$Scales[s],"_R_",
                       emis.scens$Substances[s],"_kg.xlsm") 
  
  emis.stat.long2 <- data.frame()
  for(em in Emis_routes){    
    if(em=="USE_s2"){
      Emis_data1 <- read.xlsx(Emis.loc_NR,
                              sheet = "PROB_Y_s (USE)",
                              colNames=TRUE)[,-1]
    }else if(em=="USE_w0"){
      Emis_data1 <- read.xlsx(Emis.loc_NR,
                              sheet = "PROB_Y_lw",
                              colNames=TRUE)[,-1]
    }else if(em=="USE_w1"){
      Emis_data1 <- read.xlsx(Emis.loc_NR,
                              sheet = "PROB_Y_fw",
                              colNames=TRUE)[,-1]
    }else if(em=="EOL_a_NR"){
      Emis_data1 <- read.xlsx(Emis.loc_NR,
                              sheet = "PROB_Y_a",
                              colNames=TRUE)[,-1]
    }else if(em=="EOL_a_R"){
      Emis_data1 <- read.xlsx(Emis.loc_R,
                              sheet = "PROB_Y_a",
                              colNames=TRUE)[,-1]
    }else if(em=="EOL_s2_NR"){
      Emis_data1 <- read.xlsx(Emis.loc_NR,
                              sheet = "PROB_Y_s (EOL)",
                              colNames=TRUE)[,-1]
    }else if(em=="EOL_s2_R"){
      Emis_data1 <- read.xlsx(Emis.loc_R,
                              sheet = "PROB_Y_s (EOL)",
                              colNames=TRUE)[,-1]
    }else Emis_data1 <- NA
    
    emm.mat.yr.df <- as.data.frame(Emis_data1) 
    # str(emm.mat.yr.df)
    # str(Emis_data1)
    
    ## Reshapes dataframe and puts it in long form
    data.long <- as.data.frame(emm.mat.yr.df) %>% gather(year, emmision, 1:100) 
    # str(data.long)
    data.long$year <- as.integer(as.character(data.long$year))
    
    #gets summary statistics for each year
    sum.stat <- data.long %>%   
      group_by(year) %>% 
      summarize(geo.mean = geometric.mean(emmision), P25 = quantile(emmision, 0.25,na.rm = TRUE), P75 = quantile(emmision, 0.75,na.rm = TRUE))
    
    sum.stat.long <- as.data.frame(sum.stat) %>% pivot_longer(c("P75","geo.mean", "P25"), names_to = "sum_stat", values_to = "emmision")
    sum.stat.long$scale = (emis.scens$Scales[s])
    sum.stat.long$substance = (emis.scens$Substance[s])
    sum.stat.long$emission_rout = as.factor(em)
    # str(sum.stat.long)
    
    #stores values for this cycle (compartment x scenario)
    emis.stat.long2 <- rbind(emis.stat.long2, sum.stat.long)
  }
  # str(emis.stat.long2)
  emis.stat.long3 <- rbind(emis.stat.long3, emis.stat.long2)
}

emis.stat.long3$scenario <- paste0(emis.stat.long3$scale,"_",emis.stat.long3$substance)

##Plot

str(emis.stat.long3)
unique(emis.stat.long3$scale)

LOC_As <- emis.stat.long3[emis.stat.long3$scale=="LOC",]
LOC_As$use_eol <- as.character(LOC_As$emission_rout)

LOC_As$use_eol <- sub("_s2", "",LOC_As$use_eol)
LOC_As$use_eol <- sub("_w0", "",LOC_As$use_eol)
LOC_As$use_eol <- sub("_w1", "",LOC_As$use_eol)
LOC_As$use_eol <- sub("_a", "",LOC_As$use_eol)
# LOC_As$use_eol <- substr(as.character(LOC_As$emission_rout),1,3)

LOC_As2 <- LOC_As %>%
  group_by(year,sum_stat,scale,substance,scenario,use_eol) %>%
  summarise(emmision = sum(emmision))

LOC_As3 <- LOC_As2[LOC_As2$substance=="Pb",]
p <- ggplot(subset(LOC_As3, sum_stat %in% "geo.mean"), aes(x=year, y=emmision)) +
  ylab('Emission (kg)') +
  xlab('Time (years)')+
  geom_ribbon(aes(ymin = subset(LOC_As3, sum_stat %in% "P25")$emmision, 
                  ymax = subset(LOC_As3, sum_stat %in% "P75")$emmision),
              fill = brewer.pal(7,"Blues")[2])+
  geom_line(position = "identity", alpha = 1, colour=brewer.pal(7,"Blues")[7])


p1 <- p + facet_grid(vars(scenario), vars(use_eol), scales = "free")+ theme_bw() +scale_y_continuous(trans='log10')

ggsave(filename = paste0("emission_LOCAL_useeol_Pb_LOG_",cases$paneltype[1],"_v04.jpg"),
       plot = p1,
       device = "jpeg",
       path = "figures",
       scale = 1.5,
       width = 16,
       height = 7,
       units = "cm",
       dpi = 300)
```


### Figure 4. "Verwachte blootstellingsconcentraties van zilver (Ag) , koper (Cu), en lood (Pb) afkomstig uit lood perovskiet zonnepanelen in water (wL) en bodem (sL) op lokale schaal in een scenario zonder recycling van deze metalen."

```{r Figure 4} 

# sb4n.loc <- "data/SimpleBox4DIRECT 07102021.xlsm"

#Get compartment volumes from @RISK input data
v_wL <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`v./.wL`[4:1003])
v_w1R <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`v./.w1R`[4:1003])
v_w1C <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`v./.w1C`[4:1003])
v_sL <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`v./.sL`[4:1003])
v_s2R <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`v./.s2R`[4:1003])
v_s2C <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`v./.s2C`[4:1003])
vols <- data.frame(v_wL, v_w1R, v_w1C,v_sL, v_s2R, v_s2C)

fracw_sL <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`FRACw.sL`[4:1003])
fracw_s2R <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`FRACw.s2R`[4:1003])
fracw_s2C <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`FRACw.s2C`[4:1003])
RHOSOLID <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`RHOsolid`[4:1003])
# fraca_sL <- as.double(read.xlsx(sb4n.loc,colNames=FALSE, namedRegion ="FRACa.sL") )
# fraca_s2R <- as.double(read.xlsx(sb4n.loc,colNames=FALSE, namedRegion ="FRACa.s2R"))
# fraca_s2C <- as.double( read.xlsx(sb4n.loc,colNames=FALSE, namedRegion ="FRACa.s2C"))
fracw_sdL <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`FRACw.sdL`[4:1003])
fracw_sd1R <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`FRACw.sd1R`[4:1003])
fracw_sd1C <- as.double(out.data2$`Ag_NR_Tandem`$addriskinput$`FRACw.sd1C`[4:1003])

FRACWs <- data.frame(fracw_sL, fracw_s2R, fracw_s2C)
# FRACAs <- data.frame(fraca_sL, fraca_s2R, fraca_s2C)

#initialize variables
sum.stat.long2 <- data.frame()
comp <- c("wL", "w1R", "w1C", "sL", "s2R", "s2C")
scen <- names(out.data2)

# for testing
c = "wL"
s = scen[2]

## Makes a matrix with concentrations (year x MC run) from deSolve lists, for each compartment and scenario
for(c in comp)
  for(s in scen){
    cv <- vols %>% select(matches(c))
    mat.yr <- matrix(nrow=101, ncol=1000)
    
    #new code to fill matrix, as previous one would not report on errors with DeSolve output... and skipped columns wihout valid data without warning, e.g. run 4 here.
    for(r in 1:1000) {
      
      
      #converts concentrations to ug/L for water and mg/kg for soil compartments
      if (c %in% c("wL", "w1R", "w1C")){
        try(mat.yr[,r] <- (out.data2[[s]][[r]][["SBout_kg"]][,c]/cv[r,1])*1e9/1000  ,silent = TRUE ) #divides mass by compartment volume in m3 and convert to ug/L
      } else {
        fracw <- FRACWs %>% select(matches(c))
        # fraca <- FRACAs %>% select(matches(c))
        try(mat.yr[,r] <- (f_Soil.dryweight(Conc.soil = c(out.data2[[s]][[r]][["SBout_kg"]][,c]/cv[r,1])*1e3, #*1e3 to convert from g to mg/kg dry weight, same unit as PNEC
                                            Fracs = c(1-fracw[r,1]-0.2), 
                                            RHOsolid = RHOSOLID[r] ) ) ,silent = TRUE)
        
      }
      
    }
    if(length(which(is.na(mat.yr),arr.ind = TRUE)[,2])>0){print("NA's found, please check, a Desolve SB run could have failed or a timestep failed")
      # which(is.na(mat.yr))
      which(is.na(mat.yr),arr.ind = TRUE)
      print(paste0("NA's found in Runs: " ))
      print(unique(which(is.na(mat.yr),arr.ind = TRUE)[,2]))
    }
    
    mat.yr.df <- as.data.frame(mat.yr) 
    # mat.yr.df <- mat.yr.df[-1]
    
    ## Reshapes dataframe and puts it in long form
    mat.yr.df <-t(mat.yr.df) # 1 year is 1 column
    colnames(mat.yr.df) <- seq(0, 100, by=1)   #renames columns to year number
    data.long <- as.data.frame(mat.yr.df) %>% gather(year, mass, 0:101) 
    data.long$year <- as.integer(as.character(data.long$year))
    
    #gets summary statistics for each year
    sum.stat <- data.long %>%   
      group_by(year) %>% 
      summarize(geo.mean = geometric.mean(mass), P5 = quantile(mass, 0.05,na.rm = TRUE), P95 = quantile(mass, 0.95,na.rm = TRUE))
    
    sum.stat.long <- as.data.frame(sum.stat) %>% pivot_longer(c("P95","geo.mean", "P5"), names_to = "sum_stat", values_to = "PEC")
    # sum.stat.long$Mass <- sum.stat.long$Mass
    # names(sum.stat.long)[names(sum.stat.long) == "Mass"] <- "PEC"
    sum.stat.long$scenario = s
    sum.stat.long$compartment = c
    
    #stores values for this cycle (compartment x scenario)
    sum.stat.long2 <- rbind(sum.stat.long2, sum.stat.long)
  }
}

unique(sum.stat.long2$scenario)

names(sum.stat.long2)

# dataframe with risk limits
risklimits <- expand.grid(scenario = unique(sum.stat.long2$scenario),
                     compartment = unique(sum.stat.long2$compartment))
risklimits$rlim [(risklimits$compartment == "wL" | 
              risklimits$compartment == "w1R" | 
              risklimits$compartment == "w1C") & 
             #grepl("Ag",risklimits$scenario)] <- 0.01 # Landoppervlaktewateren wettelijk JG-MKN (opgelost)
             grepl("Ag",risklimits$scenario)] <- 0.04 #(µg/L) PNEC for freshwater aquatic organisms from REACH registration dossier
risklimits$rlim [(risklimits$compartment == "wL" | 
              risklimits$compartment == "w1R" | 
              risklimits$compartment == "w1C") & 
             #grepl("Cu",risklimits$scenario)] <- 2.4  # Landoppervlaktewateren wettelijk JG-MKN (opgelost)
              grepl("Cu",risklimits$scenario)] <- 7.8 #(µg/L) PNEC for freshwater aquatic organisms from REACH registration dossier

risklimits$rlim [(risklimits$compartment == "wL" | 
              risklimits$compartment == "w1R" | 
              risklimits$compartment == "w1C") & 
              #grepl("Pb",risklimits$scenario)] <- 1.2 # Landoppervlaktewateren wettelijk JG-MKN (opgelost)
              grepl("Pb",risklimits$scenario)] <- 2.4 #(µg/L) PNEC for freshwater aquatic organisms from REACH registration dossier

risklimits$unit [(risklimits$compartment == "wL" | 
              risklimits$compartment == "w1R" | 
              risklimits$compartment == "w1C") ] <- "microgram/liter"

risklimits$rlim [(risklimits$compartment == "sL" | 
              risklimits$compartment == "s2R" | 
              risklimits$compartment == "s2C") & 
             #grepl("Ag",risklimits$scenario)] <- 1e-4 # Grond VR (droge stof)
              grepl("Ag",risklimits$scenario)] <- 1.41 #(mg/kg soil dw)PNEC for terrestrial organisms from REACH registration dossier
               
risklimits$rlim [(risklimits$compartment == "sL" | 
              risklimits$compartment == "s2R" | 
              risklimits$compartment == "s2C") & 
             #grepl("Cu",risklimits$scenario)] <- 54 # Maximale waarde bodemfunctieklasse wonen / Maximale waarden kwaliteitsklasse wonen (droge stof)
              grepl("Cu",risklimits$scenario)] <- 65 #(mg/kg soil dw)PNEC for terrestrial organisms from REACH registration dossier

risklimits$rlim [(risklimits$compartment == "sL" | 
              risklimits$compartment == "s2R" | 
              risklimits$compartment == "s2C") & 
             #grepl("Pb",risklimits$scenario)] <- 210 # Maximale waarde bodemfunctieklasse wonen / Maximale waarden kwaliteitsklasse wonen (droge stof)
              grepl("Pb",risklimits$scenario)] <- 212 #(mg/kg soil dw)PNEC for terrestrial organisms from REACH registration dossier

risklimits$unit [(risklimits$compartment == "sL" | 
              risklimits$compartment == "s2R" | 
              risklimits$compartment == "s2C") ] <- "mg/kg dw soil"


# sum.stat.Pb.Cu.Ag <- sum.stat.long2[(sum.stat.long2$scenario=="Pb_NR_Tandem" |
#                                   sum.stat.long2$scenario=="Cu_NR_Tandem"|
#                                  sum.stat.long2$scenario=="Ag_NR_Tandem")&
#                                 (sum.stat.long2$compartment=="wL"),]
# 
# risklimits.Pb.Cu.Ag <- risklimits[(risklimits$scenario=="Pb_NR_Tandem" |
#                                   risklimits$scenario=="Cu_NR_Tandem"|
#                                  risklimits$scenario=="Ag_NR_Tandem")&
#                                 (risklimits$compartment=="wL"),]

sum.stat.Ag.Cu.Pb <- sum.stat.long2[(sum.stat.long2$scenario=="Ag_NR_Tandem" |
                                  sum.stat.long2$scenario=="Cu_NR_Tandem"|
                                 sum.stat.long2$scenario=="Pb_NR_Tandem")& 
                                 (sum.stat.long2$compartment=="wL"|
                                 sum.stat.long2$compartment=="sL"),]

risklimits.Ag.Cu.Pb <- risklimits[(risklimits$scenario=="Ag_NR_Tandem" |
                                  risklimits$scenario=="Cu_NR_Tandem"|
                                 risklimits$scenario=="Pb_NR_Tandem")&
                                 (risklimits$compartment=="wL"|
                                 risklimits$compartment=="sL"),]

#make figure 4
p <- ggplot(subset(sum.stat.Ag.Cu.Pb, sum_stat %in% "geo.mean"), aes(x=year, y=PEC)) +
  ylab('Concentration (ug/L or mg/kg dw)') +
  xlab('Time (years)')+
  geom_ribbon(aes(ymin = subset(sum.stat.Ag.Cu.Pb, sum_stat %in% "P5")$PEC, 
                  ymax = subset(sum.stat.Ag.Cu.Pb, sum_stat %in% "P95")$PEC),
              fill = brewer.pal(7,"Greens")[2])+
  geom_line(position = "identity", alpha = 1, colour=brewer.pal(7,"Greens")[7])

p1 <- p + geom_hline(data = risklimits.Ag.Cu.Pb, aes(yintercept = rlim))

p3 <- p1 + facet_grid(vars(scenario),vars(compartment), scales = "free")+ theme_bw() +scale_y_continuous(trans='log10')+ylab('LOG-Concentration (ug/L or mg/kg dw)') 

ggsave(filename = paste0("PEC_Ag_Cu_Pb_RLIM_v04.jpg"),
       plot = last_plot(),
       device = "jpeg",
       path = "figures",
       scale = 2,
       width = 16,
       height = 10,
       units = "cm",
       dpi = 300)

